---
title: "Writing a CV"
subtitle: "Producing and formatting entry listings"
engine: markdown
---

A CV built with Quarto vitae is an ordinary `.qmd` document. Its sections are
plain headings, and each listing — education, employment, awards — is a **table**
plus a **style** saying how it should look.

Where the table comes from is up to you. Write it by hand, generate it from R,
Python or Julia, or read it from a file. The result is the same, so you can start
by typing a table and move to generated data later without rewriting anything.

## The shortest possible listing

```markdown
---
format: awesomecv-pdf
---

# Education

::: {.entries style="detailed"}
| what             | with                | when      | where         |
|------------------|---------------------|-----------|---------------|
| Ph.D. in Physics | University of Paris | 1891–1895 | Paris, France |
:::
```

The `.entries` div marks the table as a CV listing, and `style` picks the layout.
Column names are passed straight to the template, so a table headed with the
fields your template expects needs no configuration at all.

::: {.callout-tip}
## Which styles exist?

That depends on your template. Most provide `detailed` (a multi-line entry with
a bulleted description), `brief` (one line each), and often `skills`. Ask for one
that does not exist and the filter tells you what is available.
:::

## Naming your own columns

Real data rarely has columns called `what` and `when`. Rename them with `fields`,
which reads like `dplyr::rename()` — the new name first:

```markdown
::: {.entries style="detailed" fields="{what: Degree, with: Institution, when: Years}"}
| Degree           | Institution         | Years     |
|------------------|---------------------|-----------|
| Ph.D. in Physics | University of Paris | 1891–1895 |
:::
```

Only the columns you name are renamed; everything else keeps its own heading. So
`fields` is a short list of corrections, not a full schema.

### The conventional fields

Templates agree on five names, which is what lets one table feed several CV
designs:

| Field   | Holds                                              |
|---------|----------------------------------------------------|
| `what`  | the thing itself — degree, job title, award        |
| `when`  | the date or range                                  |
| `with`  | the associated body — university, employer, funder |
| `where` | the place                                          |
| `why`   | description or achievements, often several items   |

They are intentionally vague so the same five work for education, employment and
awards alike. A template may support others; check its documentation.

## Descriptions with several items

`why` usually wants more than one line. There are three ways to write it.

**A grid table**, whose cells can contain a bullet list:

```markdown
::: {.entries style="detailed"}
+-----------+---------------------+-----------+-----------------------------+
| what      | with                | when      | why                         |
+===========+=====================+===========+=============================+
| Professor | University of Paris | 1906–1934 | - First woman to hold the   |
|           |                     |           |   chair of General Physics  |
|           |                     |           | - Directed the Curie        |
|           |                     |           |   Laboratory                |
+-----------+---------------------+-----------+-----------------------------+
:::
```

**A pipe table with `collapse`**, writing one row per item. Consecutive rows that
differ *only* in the named column are merged:

```markdown
::: {.entries style="detailed" collapse="why"}
| what      | with                | when      | why                        |
|-----------|---------------------|-----------|----------------------------|
| Professor | University of Paris | 1906–1934 | First woman to hold the chair |
| Professor | University of Paris | 1906–1934 | Directed the Curie Laboratory |
:::
```

**Several columns at once**, by mapping them onto one field. Empty cells are
skipped:

```markdown
::: {.entries style="detailed" fields="{why: [achievements, honour]}"}
```

## Generating listings from code

Put the same options in a code cell, under `vitae`. The cell's table output is
picked up wherever it appears:

````markdown
```{r}
#| echo: false
#| vitae:
#|   style: brief
#|   fields:
#|     what: prize
#|     with: awarded_by
#|     when: year
awards <- read.csv("awards.csv")
awards[order(-awards$year), ]
```
````

This is the real reason to generate listings: **filtering and sorting happen in
the language you already use.** There is no CV-specific query syntax to learn —
use `dplyr`, `polars`, or plain indexing.

Note the cell simply names the data frame. Most CV templates set `df-print: kable`
so a printed data frame becomes a table automatically, and the Jupyter engine
does the same for a pandas `DataFrame`:

````markdown
```{python}
#| echo: false
#| vitae: {style: brief}
import pandas as pd
pd.DataFrame({"what": ["Nobel Prize"], "when": [1903]})
```
````

Use `echo: false` unless you want the code itself in your CV. The cell and its
output are replaced by the formatted listing either way.

### When the output is not a table

Auto-printing only works for objects your engine knows how to render — a data
frame or a pandas `DataFrame`, but not a plain list or dictionary. Anything else
prints verbatim, and you get a warning:

```
(W) vitae: code cell (style 'brief') produced no table, so no entries were rendered.
```

## Reading CSV directly

For a path that does not depend on the engine at all, print CSV and say so:

````markdown
```{r}
#| echo: false
#| vitae:
#|   style: brief
#|   input: csv
write.csv(awards, stdout(), row.names = FALSE)
```
````

This works from any language — printing text is something they all do:

````markdown
```{python}
#| echo: false
#| vitae: {style: brief, input: csv}
import sys
awards.to_csv(sys.stdout, index=False)
```
````

`input: csv` is worth reaching for when:

- your engine will not render the object as a table
- values contain **commas, quotes or line breaks**, which a markdown table cannot
  carry
- you want output that does not change when a library updates its printing

Reading a `.csv` file needs no code beyond loading it:

````markdown
```{r}
#| echo: false
#| vitae: {style: detailed, input: csv}
writeLines(readLines("cv-data/employment.csv"))
```
````

## Formatting entry text

Cell contents are ordinary markdown, converted correctly for whatever you are
rendering to. This works in every input method above.

| You write            | You get                        |
|----------------------|--------------------------------|
| `**Solvay Council**` | bold                           |
| `*Sorbonne*`         | italic                         |
| `[report](https://…)`| a link                         |
| `$E = mc^2$`         | maths                          |
| `100%`, `AT&T`       | literal — escaping is automatic |

```markdown
::: {.entries style="brief"}
| what                                         | when |
|----------------------------------------------|------|
| Member, **Solvay Council**                   | 1911 |
| Director, Red Cross Radiology (100% wartime) | 1914 |
| See [the report](https://example.org) & more | 1920 |
:::
```

You never need to escape LaTeX specials yourself. `%`, `&`, `_` and `#` are
handled, and doing it manually will double-escape.

::: {.callout-warning}
## Line breaks in pipe tables

A pipe table cell cannot contain a line break, and an unescaped `|` will break
the row. If your text needs either, use a grid table or `input: csv`.
:::

## Several listings at once

One `.entries` div may hold more than one table, and the options apply to all of
them. Useful when a section has natural groupings:

```markdown
::: {.entries style="skills"}
| what       | with                             |
|------------|----------------------------------|
| Techniques | Fractional crystallisation       |

| what       | with                             |
|------------|----------------------------------|
| Languages  | Polish, French, Russian          |
:::
```

The same applies to a code cell that emits several tables in a loop.

## Overriding a template

To change how one style looks without leaving your document, point it at your own
template file:

```yaml
---
format: awesomecv-pdf
vitae:
  templates:
    timeline: templates/timeline.tex
---
```

Any style can be overridden or added this way, including one your CV template
does not provide. See [Creating CV templates](creating-templates.qmd) for how to
write the file.

## Options reference

Written as div attributes, or under `vitae:` in cell options.

| Option     | Default    | Meaning                                              |
|------------|------------|------------------------------------------------------|
| `style`    | `detailed` | Which layout to use                                  |
| `fields`   | identity   | Rename columns, `{what: Degree}`                     |
| `collapse` | off        | Merge consecutive rows differing only in this column |
| `input`    | off        | Read printed data instead of a table (`csv`)         |

Both spellings carry the same value. On a div it is a flow mapping:

```markdown
::: {.entries style="brief" fields="{what: prize, when: year}"}
```

In cell options it may be a nested block:

````markdown
```{r}
#| vitae:
#|   style: brief
#|   fields:
#|     what: prize
#|     when: year
```
````

## Troubleshooting

`produced no table, so no entries were rendered`
: The engine printed your object verbatim. Use `input: csv`, or convert it
  yourself with `knitr::kable()` or `.to_markdown()`.

`unknown style 'x'; this format provides …`
: A typo, or a style your template does not implement. The listed names are what
  it supports.

`could not parse fields '…'`
: `fields` needs a mapping in braces: `{what: Degree}`, not `what=Degree`.

`ignoring unknown attribute 'x' on .entries div`
: A misspelled option name. Check it against the table above.

**Entries render but a column is missing.** The template only shows the fields it
knows about. Check the column reached it under the right name — an unrenamed
`Degree` stays `Degree`, and a template looking for `what` will not find it.
